---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
---
First of all, we need to unzip and read the input file with R code:
```{r}
unzip("activity.zip")
activityRecords<-read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
The total number of steps taken per dat is calculate:
```{r}
steps.date<-aggregate(steps~date,data=activityRecords,FUN=sum)
```

To understant the data, by following code, the histogram of daily step sums is produced:
```{r}
barplot(steps.date$steps, names.arg=steps.date$date, xlab="Days", ylab="NumOfsteps")
```

Mean and median is calculated by following code:
```{r}
mean(steps.date$steps)
median(steps.date$steps)
```

## What is the average daily activity pattern?
With 5-minute intervals, average step numbers are summarized as a plot:
```{r}
steps.interval <- aggregate(steps~interval,data=activityRecords,FUN=mean)
plot(steps.interval,type="l")
```
Max step numbers are calculated:
```{r}
steps.interval$interval[which.max(steps.interval$steps)]
```

## Imputing missing values
Total number of missing values are calculated:
```{r}
sum(is.na(activityRecords$steps))
```

Devise a strategy: Actualy, if there is no data, I prefer to replace with means.

Missing values are replaced:
```{r}
cleanDataSet <- merge(activityRecords, steps.interval, by="interval", suffixes=c("",".y"))
nas <- is.na(cleanDataSet$steps)
cleanDataSet$steps[nas] <- cleanDataSet$steps.y[nas]
cleanDataSet <- cleanDataSet[,c(1:3)]
```
You can find the related plots:
```{r}
steps.date<-aggregate(steps~date,data=cleanDataSet,FUN=sum)
barplot(steps.date$steps,names.arg=steps.date$date,xlab="Days",ylab="NumOfsteps")
mean(steps.date$steps)
median(steps.date$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
Day type variable is created:
```{r, cache=TRUE}
dType <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "Weekend"
    } else {
        "Week Day"
    }
}
cleanDataSet$dType <- as.factor(sapply(cleanDataSet$date, dType))
```

Plots are produced:
```{r}
par(mfrow=c(2,1))
for (type in c("Weekend", "Week Day")) {
 steps.type <- aggregate(steps ~ interval,data=cleanDataSet,subset=cleanDataSet$dType==type,FUN=mean)
 plot(steps.type, type="l", main=type)
}
```
